Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-17 paper/paper.tex

\subsection{Static Provisioning Dynamic Scheduling (SPDS)}

This is the simplest strategy based on static provisioning of resources and
on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
together with the hourly price $p$ of a VM it is possible to compute the number
of VMs $N_{VM}$ to provision so that they consume the whole budget between the
deadline.

@@

Although simple, such provisioning plan has the advantage that it
minimizes the number of provisioning and deprovisioning requests.
Once the resources are provisioned, the tasks are scheduled to VMs using dynamic
priority-based algorithm shown in Algorithm~\ref{alg:ds}.

% This is an example of how we can do algorithms
\begin{algorithm}
\caption{Priority-based scheduling}
\label{alg:ds}
\begin{algorithmic}[1]
\Procedure{schedule}{}
    \State $Q\gets set\ of\ released\ tasks$
    \State $P\gets empty\ priority\ queue$
    \While{$Q \neq \emptyset$}
    	\For{$task\ t\in Q$}
    		\State \Call{Insert}{$t,P$} 
    	\EndFor
    	\While{$FreeVMs \neq \emptyset \wedge P \neq \emptyset $}
    		\State $VM\gets$ \Call{SelectRandom}{$FreeVMs$}
    		\State $t\gets$ \Call{Pop}{$P$}
    		\State \Call{Submit}{$t,VM$}
    	\EndWhile
    	\State \Call{Wait}{$task\ finished$}
    	\State \Call{Update}{$Q$}
    \EndWhile
\EndProcedure
\end{algorithmic} 
\end{algorithm}

Initially all the
released tasks from all the workflows are submitted to the queue $Q$. Then, they
are inserted into priority queue $P$ based on the priority of workflows. If
there are free VMs available, the tasks from the priority queue are submitted to
the randomly choosen VMs. The process is repeated each time a task finishes.
Using the priority queue guarantees that tasks from lower priority workflows are
always submitted later than the higher priority ones and that the lower priority
tasks can occupy idle VMs. However, as there is no preemption, long-running
low-priority tasks may delay execution of high-priority tasks when such ones
become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
SPSS algorithm.

@@

>>>>

\subsection{Static Provisioning Dynamic Scheduling (SPDS)}
\label{sec:spds}
This is the simplest strategy based on static provisioning of resources and
on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
together with the hourly price $p$ of a VM it is possible to compute the number
of VMs $N_{VM}$ to provision so that they consume the whole budget before the
deadline.

@@

Although simple, such provisioning plan has the advantage that it
minimizes the number of provisioning and deprovisioning requests.
Once the resources are provisioned, the tasks are scheduled to VMs using dynamic
priority-based Algorithm~\ref{alg:ds}.

\begin{algorithm}
\caption{Priority-based scheduling}
\label{alg:ds}
\begin{algorithmic}[1]
\Procedure{schedule}{}
    \State $Q\gets set\ of\ released\ tasks$
    \State $P\gets empty\ priority\ queue$
    \While{$Q \neq \emptyset$}
    	\For{$task\ t\in Q$}
    		\State \Call{Insert}{$t,P$} 
    	\EndFor
    	\While{$FreeVMs \neq \emptyset \wedge P \neq \emptyset $}
    		\State $VM\gets$ \Call{SelectRandom}{$FreeVMs$}
    		\State $t\gets$ \Call{Pop}{$P$}
    		\State \Call{Submit}{$t,VM$}
    	\EndWhile
    	\State \Call{Wait}{$task\ finished$}
    	\State \Call{Update}{$Q$}
    \EndWhile
\EndProcedure
\end{algorithmic} 
\end{algorithm}

Initially all the
released tasks from all the workflows are submitted to the queue $Q$. Then, they
are inserted into priority queue $P$ based on the priority of workflows. If
there are free VMs available, the tasks from the priority queue are submitted to
the randomly choosen VMs. The process is repeated each time a task finishes.
Using the priority queue guarantees that tasks from lower priority workflows are
always submitted later than the higher priority ones and that the lower priority
tasks can occupy idle VMs. However, as there is no preemption, long-running
low-priority tasks may delay execution of high-priority tasks when such ones
become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
SPSS algorithm. It can be seen that tasks from lower priority workflows can fill
idle time of provisioned virtual machines.

@@

Static provisioning plan used in SPDS algorithm allows no changes to resource
pool at runtime. However, such plan may not be optimal when resource demand
changes during execution of workflows. In policy-based provisioning
(or auto-scaling) systems the metrics such as resource
utilization or queue length are typicaly used to estimate curret resource
demand. Our DPDS algorithm is based on resource utilization: the provisioner
module periodically computes utilization as the percentage of idle VMs. The
provisioner algorithm detects which VMs are finishing their hourly billing cycle
during the next provisioning period and decides whether they should continue
running, should be terminated or new VMs should be provisioned: see
Algorithm~\ref{alg:prov}.

\begin{algorithm}
\caption{Dynamic provisioner}
\label{alg:prov}
\begin{algorithmic}[1]
\Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
$t$: current time; $u_h$: upper utilization thershold; $u_l$: lower utilization
threshold.
\Procedure{provision}{}
    \State $V_C\gets set\ of\ VMs\ completing\ billing\ cycle$
    \State $n_T\gets\ 0$ \Comment{number of VMs to terminate} 
    \If{$ b - c < p * |V_C| \vee t > d $ }
    	\State $V_R\gets set\ of\ running\ VMs$
    	\State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$ 
    	\State $V_T\gets select\ n_T VMs\ to\ terminate\ from\ V_C$
    	\State \Call{Terminate}{$V_T$}
    \Else 
    	\If{$u>u_h$}
    		\State \Call{Start}{$new\ VM$}
    	\EndIf
    	\If{$u<u_l$}
    		\State $V_F\gets free\ (idle)\ VMs$
    		\State $n_T\gets \lceil|V_F|/2\rceil$ 
    		\State \Call{Terminate}{$V_T$}
    	\EndIf 
    \EndIf
\EndProcedure
\end{algorithmic} 
\end{algorithm}

The number of instances to terminate $n_T$ computed in line 6 is the number that
would overrun the budget if not terminated. On the other hand $n_T$ computed in
line 15 prevents terminating all the instances too quickly, but makes sure that
if there is only one instance left it should be terminated. However, the
termination process is not immediate: to make sure that VMs that have been
already paid for are not terminated prematurely the {\em Terminate()} procedure
terminates only these instances that are close to the end of their billing
cycle. The remaining time threshold is calculated as a sum of provisioner
interval and estimated termination delay. This guarantees that no VMs can
overrun the assigned budget. To avoid uncontrolled increase of number of
instances, which may happen in the case of highly parallel workflows,
provisioner checks also the maximum scaling factor from the initial number of
provisioned VMs and prohibits starting new ones when this threshold is reached.
